Inferring disease architecture and predictive ability\\with LDpred2-auto

### 2022-08-19 paper/paper-infer.tex
save

@@

The HapMap3 variants generally provide a good coverage of the whole genome.
We recall that the set of 1,054,330 HapMap3 variants recommended to use for LDpred2 \cite[]{prive2020ldpred2} is a subset of the original set of HapMap3 variants, which does not include duplicated positions (e.g.\ multi-allelic variants), ambiguous variants (e.g.\ both 'A' and 'T' or 'C' and 'G'), and which includes SNPs only (e.g.\ no indel).
Here we propose to extend the set we have used for LDpred2 until then. 
This extension aims at making sure many genetic variants are well tagged by the extended set, which we call HapMap3+.
To design this new set, we first read all variants from the UK Biobank (UKBB) with a minor allele frequency (MAF) larger than 0.005 in the whole data (i.e.\ the one in the MFI files). 
We then compute all pairwise correlations between variants within a 1 Mb distance, restricting to squared correlations larger than 0.3, and using all unrelated UKBB individuals excluding all White British (Field 22006) to have a diverse set of populations.
Finally, we design an algorithm which aims at maximizing the tagging of previously read variants. We want to maximize $\sum_j \max_{k \in \text{HapMap3+}} r_{j,k}^2$, where $j$ spans the whole set of variants read, while $k$ spans the variants kept in the new set.
We start by including all previously used HapMap3 variants.
For the sake of simplicity, we then use a greedy approach, where we repeatedly include the variant which increases this sum most, until no variant improves it by more than 2. 
Note that we allow only non-ambiguous SNPs to be included.
This results in an extended set of 1,444,196 SNPs.

>>>>

\subsection*{Data for the UK Biobank analyses}

We use the set of 1,054,330 HapMap3 variants recommended to use for LDpred2 \cite[]{prive2020ldpred2}, and the same 356,409 individuals of Northwestern European ancestry as in the simulations.
As covariates, we first recompute PCs for this subset of individuals using function \texttt{snp\_autoSVD} from R package bigsnpr and keep four PCs based on visual inspection \cite[]{prive2017efficient,prive2020efficient}. We also use sex (Field 22001), age (Field 21022), birth date (combining Fields 34 and 52) and deprivation index (Field 189) as additional covariates.
We randomly sample 50,000 individuals to form a test set (to evaluate the predictive models), and use the remaining individuals to form a training set (to run the GWAS).

We use the LD matrix with independent LD blocks computed in \cite{prive2021identifying}.
We design two other LD matrices: one using a subset of 2000 individuals from the previously selected ones (which we call ``hm3\_small''), and one based on 10,000 individuals from around South Europe by using the ``Italy'' center defined in \cite{prive2021high} (``hm3\_altpop'').
We apply the optimal algorithm developed in \cite{prive2021optimal} to obtain independent LD blocks, as recommended in \cite{prive2021identifying}.
We finally define a fourth LD reference by extending the set of HapMap3 variants (see next Methods section) and using 20,000 individuals from the previously selected ones (``hm3\_plus'').

The HapMap3 variants generally provide a good coverage of the whole genome.
We recall that the set of 1,054,330 HapMap3 variants recommended to use for LDpred2 \cite[]{prive2020ldpred2} is a subset of the original set of HapMap3 variants, which does not include duplicated positions (e.g.\ multi-allelic variants), ambiguous variants (e.g.\ both 'A' and 'T' or 'C' and 'G'), and which includes SNPs only (e.g.\ no indel).
Here we propose to extend the set we have used for LDpred2 until then. 
This extension aims at making sure many genetic variants are well tagged by the extended set, which we call HapMap3+.
To design this new set, we first read all variants from the UK Biobank (UKBB) with a minor allele frequency (MAF) larger than 0.005 in the whole data (i.e.\ the one in the MFI files). 
We then compute all pairwise correlations between variants within a 1 Mb distance, restricting to squared correlations larger than 0.3, and using all unrelated UKBB individuals excluding all White British (Field 22006) to have a diverse set of populations.
Finally, we design an algorithm which aims at maximizing the tagging of previously read variants. We want to maximize $\sum_j \max_{k \in \text{HapMap3+}} r_{j,k}^2$, where $j$ spans the whole set of variants read, while $k$ spans the variants kept in the new set.
We start by including all previously used HapMap3 variants.
For the sake of simplicity, we then use a greedy approach, where we repeatedly include the variant which increases this sum most, until no variant improves it by more than 2. 
Note that we only allow non-ambiguous SNPs to be included.
This results in an extended set of 1,444,196 SNPs, of which we compute the correlation between variants and apply the optimal algorithm developed in \cite{prive2021optimal} to obtain 431 independent LD blocks.

@@
